---
title: "Assignment 1"
subtitle: "Web scraping class"
author: "Viktória Kónya"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:

  prettydoc::html_pretty:
    highlight: github
    toc: true
    theme: cayman
---

**Scraped news portal** : https://www.portfolio.hu/

```{r warning = F, message = F, echo = F}

# Import packages
library(xml2) 
library(rvest)
library(data.table)
library(kableExtra)
library(dplyr)

```



## Scraping function
I needed to add an if-else statement as the first page of each topic (page=1) has different structure than the rest of the pages. In addition, the first page has 3 headline articles which do not have subtitles and needed to appended to the the rest of the articles on the page.

```{r}

get_one_page <- function(url) {
  
  # Concatenate url
  # url <- paste0(main_url, rovat, "?page=", as.character(page_num))
  

  # Different logic is needed for the first and the rest of the pages
  
  if (grepl( "page=1", url) == TRUE) {
    
    
    ### First page 
    
    # Get link
    t <- read_html(url) 
    
    
    # A. Get data from articles 
    
    # Titles
    titles_articles <- t %>% html_nodes('.category-articles') %>% html_nodes('.art') %>% html_nodes('a') %>% html_text() 
    
    # Subitles
    subtitles_articles <- t %>% html_nodes('.category-articles') %>% html_nodes('.art') %>% html_nodes('p') %>% html_text() 
    
    # Links
    links_articles <- t %>% html_nodes('.category-articles') %>% html_nodes('.art') %>% html_nodes('a') %>%  html_attr('href')
    links_articles <- links_articles[grepl( paste0("https://www.portfolio.hu/" ,rovat,"/"), links_articles)]
    
    # Put articles part together
    df_articles <- data.frame('titles' = titles_articles, 'subtitles' = subtitles_articles, 'links' = links_articles)
    
    
    # B. Get data from headline
    
    # Titles
    titles_headline <- t %>% html_nodes('.category-articles') %>%  html_nodes('.overlay')  %>% html_nodes('a') %>%  html_text() 
    titles_headline <- titles_headline[seq(2, length(titles_headline), by = 2)]
    
    # Links
    links_headline <- t %>% html_nodes('.category-articles') %>%  html_nodes('.overlay')  %>% html_nodes('a') %>% html_attr('href')
    links_headline <- links_headline[seq(2, length(links_headline), by = 2)]
    
    # Put header part together
    df_headline <- data.frame('titles' = titles_headline, 'subtitles' = NA, 'links' = links_headline)
    
    
    # C. Append two parts
    df <- rbind(df_headline, df_articles)
    
    return(df)
    
    
  }else{
    
    ### Rest of the pages 
    
    # Get link
    t <- read_html(url) 
    
    # Tites
    titles <- t %>% html_nodes('.article-lists') %>% html_nodes('h3') %>% html_text() 
    
    # Subtites
    subtitles <- t %>% html_nodes('.article-lists') %>% html_nodes('p') %>% html_text() 
    
    # Link
    links <- t %>% html_nodes('.article-lists') %>% html_nodes('a') %>%  html_attr('href')
    links <- links[grepl( paste0("https://www.portfolio.hu/",rovat,"/"), links)]
    
    df <- data.frame('titles' = titles, 'subtitles' = subtitles, 'links' = links)
    return(df)
  
    
  } # end else
  
  
} # end func

```


## Scrape one page
We can change the name of the topic in the url.
```{r}

# Scrape one page
main_url <- "https://www.portfolio.hu/"

rovat <- "gazdasag"
page_num <- 1
df1 <- get_one_page(paste0(main_url, rovat, "?page=", as.character(page_num)))

kable(head(df1)) %>%
  kable_styling(full_width = F, font_size = 11)

page_num <- 2
df2 <- get_one_page(paste0(main_url, rovat, "?page=", as.character(page_num)))

kable(head(df2)) %>%
  kable_styling(full_width = F, font_size = 11)

page_num <- 1
rovat <- "deviza"
df3 <- get_one_page(paste0(main_url, rovat, "?page=", as.character(page_num)))

kable(head(df3)) %>%
  kable_styling(full_width = F, font_size = 11)

page_num <- 5
rovat <- "befektetes"
df4 <- get_one_page(paste0(main_url, rovat, "?page=", as.character(page_num)))

kable(head(df4)) %>%
  kable_styling(full_width = F, font_size = 11)


```

## Scrape multiple pages
```{r}

# Multiple pages
rovat <- "gazdasag"
page_num <- 1

links <- paste0(paste0(main_url, rovat, "?page=", 1:5))

list_of_dfs <- lapply(links, get_one_page)

list_of_dfs_appended <- rbindlist(list_of_dfs)

kable(head(list_of_dfs_appended)) %>%
  kable_styling(full_width = F, font_size = 11)

```

