---
title: "Assignment 2"
subtitle: "Web scraping class"
author: "Viktória Kónya"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:

  prettydoc::html_pretty:
    highlight: github
    toc: true
    theme: cayman

---

**The Movie Database (TMDb) API Documentation** : https://developers.themoviedb.org/3/getting-started/introduction

```{r warning = F, message = F, echo = F}

# Import packages
library(rvest)
library(data.table)
library(jsonlite)
library(httr)
library(ggplot2)
library(tidyverse)
library(kableExtra)

```

```{r global_options, echo = FALSE, include = FALSE}
options(width = 999)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,
                      cache = FALSE, tidy = FALSE, size = "small")
```

### Import file with TMDb movie_ids
In order to make GET requests, the movie_id of the movies is needed. 
The list of all movie_ids can be downloaded from the TMDB API site in a zipped file. I unzipped and added this JSON file to my GitHub repo, made a subset of that and I will call it from my GitHub.

See desciption here: https://developers.themoviedb.org/3/getting-started/daily-file-exports

```{r, echo = F}
# DO NOT EXECUTE THIS !!!
# movie_ids <- stream_in(file("https://raw.githubusercontent.com/viktoriakonya/ECBS-5306-Coding-2-Web-Scraping-with-R/master/Assignment_2/data/movie_ids_11_16_2021_downloaded.json"))
# movie_ids <- movie_ids %>% arrange(desc(popularity)) 
# movie_ids_short <- movie_ids %>% head(100)
# stream_out(movie_ids_short, file("movie_ids_short.json"))
```

### Import JSON file with movie_ids
```{r, error = F, warning = F, messages = F}

movie_ids_short <- stream_in(file("https://raw.githubusercontent.com/viktoriakonya/ECBS-5306-Coding-2-Web-Scraping-with-R/master/Assignment_2/data/movie_ids_short.json"))

kable(head(movie_ids_short)) %>%
  kable_styling(full_width = F, font_size = 11)
```


### GET APIs of 100 movies
```{r}


# Get APIs
get_api <- function(url) {
  
  t <- GET(url)
  t <- fromJSON(content(t, "text"))
  
  # ID
  id <- t$id
  
  # Production companies
  production_companies <- paste0(t$production_companies$name, collapse = ", ")
  
  # Budget
  runtime <- t$runtime
  
  # Genres
  genres <- paste0(t$genres$name, collapse = ", ")

  # Title
  title <- t$original_title
  
  # Overview
  overview <- t$overview
  
  # Budget
  budget <- t$budget
  
  # Revenue
  revenue <- t$revenue
  
  # Popularity
  popularity <- t$popularity
  
  # Vote Average
  vote_average <- t$vote_average
  
  # Vote Count
  vote_count <- t$vote_count
  
  # Df
  df <- data.frame(id, 
                   title, 
                   production_companies, 
                   runtime, budget, 
                   revenue, 
                   popularity, 
                   vote_average, 
                   vote_count, 
                   overview)
  
  colnames(df) <- c("ID", 
                    "Title", 
                    "Production companies",  
                    "Runtime", 
                    "Budget", 
                    "Revenue",
                    "Popularity", 
                    "Vote Average", 
                    "Vote Count", 
                    "Overview")
  
  return(df)
  
}

# Define API key and movie_id list
api_key <- '1858116a97594edd8c00b69df8d51d3f'
movie_ids <- movie_ids_short$id

# Concatenate links
links <- paste0("https://api.themoviedb.org/3/movie/",movie_ids,"?api_key=", api_key)

# Create Df
list_of_dfs <- lapply(links, get_api)
list_of_dfs_appended <- rbindlist(list_of_dfs)

kable(head(list_of_dfs_appended)) %>%
  kable_styling(font_size = 11) 

```

### Visualizations

#### Distribution of movie runtimes
```{r, echo = F}

list_of_dfs_appended %>%
 filter(Runtime > 0) %>%
 ggplot() +
 aes(x = Runtime) +
 geom_density(adjust = 1, fill = "#41ab5d", alpha = 0.8) +
  labs(title = "Distribution of movie runtimes", y = "Density", x = "Runtime") +
 theme_bw() +
 theme(
    plot.title = element_text(size = 14L,
    face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    legend.position = "none"
 )

```

#### Top 10 movies with highest revenues
```{r, echo = F}

list_of_dfs_appended %>% 
  arrange(desc(Revenue)) %>% 
  mutate(Revenue_MM = Revenue / 1000000 )  %>% 
  head(10) %>% 
  ggplot() +
  aes(x = reorder(Title, Revenue_MM), weight = Revenue_MM ) +
  scale_fill_viridis_c(begin = 0.2, end = 0.7) +
  geom_bar( aes(fill = -Revenue_MM)) +
  labs(title = "Top 10 movies with highest revenues", y = "Revenue (million $)", x = NULL) +
  coord_flip() +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14L,
    face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    legend.position = "none"
  )

```

#### Relationship between vote count and average vote
```{r, echo = F}
list_of_dfs_appended %>% 
  filter(`Vote Count` >= 200) %>% 
  filter(`Vote Average` >= 5) %>% 
  ggplot() +
  aes(
    x = `Vote Count`,
    y = `Vote Average`,
    colour = `Vote Average`
  ) +
  geom_point(shape = "circle", size = 1.5) +
  geom_smooth(method='lm', se = F, color = 'gray20') +
  scale_color_viridis_c(option = "viridis", direction = -1, begin = 0.2, end = 0.8) +
  labs(title = "Relationship between vote count and average vote") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14L,
    face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    legend.position = "none"
  )

```

